### 2014-09-06 relax1.tex
relax I typos fixed

\subsection{SSA Form of the Trace}~\label{sec:ssa}
First of all, we rewrite the trace into SSA form, such that every variable is
defined exactly once in the trace. This requirement is a prerequisite for 
the computation of the mapping, in which each variable is mapped to exactly 
one value. Another side effect is that the def/use chains become explicit 
in SSA form, which simplifies the following analysis steps.

\tool\ handles local assignments and heap accesses differently.
\begin{itemize}
\item {\bf Local Assignment\ } The SSA form for local assignments 
resembles the static SSA form in compiler optimizations, 
except that loops and recursions are fully resolved in the trace, 
obviating the need for {\sf Phi} nodes\footnote{While \tool in fact uses
Phi nodes in the second relaxation, for simplicity during discussion, we 
do not assume Phi nodes in this stage.} . More concretely, we replace the 
variable $v$ defined in an event, as well as the following uses of the 
definition, to a new variable $v^{id}$, where $id$ is the unique id 
of the event. The uniqueness of the id guarantees that no two events 
define the same variable, i.e., each variable is defined exactly once.  
In Figure~\ref{fig:running2}, for example,
the local variable $x$ defined at line 4 is renamed as $x^4$.
% where the line number is also used as event id,  the local variable $x$ defined at line 4 is renamed as $x^4$ in the trace, while $x$ defined at line 9 is renamed as $x^9$, which is used in the following two events.  At line 5, where the object $o3$ is created,  we define a reference variable $w^5$.  Internally,  the object $o3$ is encoded as a unique integer that indicates its heap address.
\item {\bf Heap accesses\ } The def/use relation between local heap accesses 
can be determined locally. Therefore, we treat them in the same way as 
local assignments. The def/use relations between shared heap accesses 
are more complex. We may link a read with different writes under different 
schedules, e.g., the read of {\tt y} at line 4 in Figure \ref{fig:running} can 
be from before line 1, line 1, or line 3. Therefore, we introduce two symbols to represent the shared read and shared write respectively and use them 
to reason about the possible options.
%establish different links flexibly. 
The details are as follows. 
\begin{itemize}
\item {\bf Local Heap Accesses\ } Given the write $x.f=y$ to a local heap 
location at event $e_{id}$,  we introduce a fresh local variable 
{\tt $l^{id}_{o.f}$} ($o$ is the runtime object referenced by $x$) to 
replace $x.f$. We also replace the uses of the location $o.f$ accordingly.
Here  the subscript $o.f$ is interpreted as a heap location, i.e., the 
field $f$ of the object $o$. Consider the example in 
Figure~\ref{fig:running2}. At lines 6-7, we replace the local heap 
accesses as the local variable $l^6_{o3.f}$.
%An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
\item  {\bf Shared Heap Accesses\ }  Given the write $x.f=y$ or the read $y=x.f$ of a shared location at event $e_{id}$, we introduce two symbols, $W^{id}_{\tt o.f }$ and $R^{id}_{\tt  o.f }$ respectively, where $o$ is the runtime object referenced by $x$. Similarly, the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$. Note that in local heap encoding, 
both read and write are bounded to the same variable, whereas in shared
heap encoding, we use different variables to denote read and write, allowing
the solver to reason about the different dataflow.
\end{itemize}
We use the heap location such as $o.f$ in the symbolic form of heap accesses.
 An underlying assumption is that the heap location remains unchanged, 
which we ensure through additional constraints (Section~\ref{sec:relax1}).
\end{itemize}

@@

\begin{itemize}
\item {\bf Control Flow Constraints\ } The branches in the predicted run 
should take the same decisions as in the original trace so that each thread 
reproduces the same sequence of events\footnote{We allow path differences in
the second relaxation. This is to avoid presenting the encoding
of both relaxations together, which is overly complex.} 
%the encodings of both relaxation together, .  
%Specifically, we only need to reason about the branches prior to $e_i$.

Without loss of generality, we assume the branch event $e_k$ is in the form of $if(x<y)$. Then  we require that 
$$
\bigwedge_{e_k \in prefix(\tau,e_i) \wedge  inst^{e_k}=if(x<y).\ }  \theta'(x)<\theta'(y) 	\equiv \theta(x)<\theta(y)
$$ 
The constraint specifies that the branch condition should be evaluated as 
the same boolean value in the predicted run $\tau'$ as in the original 
trace $\tau$.
 Importantly, Unlike existing analyses \cite{yannis,pldi14}, we do not 
pose the requirement that the values flowing into branching statements 
remain the same, but adopt the relaxed requirement that 
the branch outcome remains the same.
%the evaluation 
%of branching expressions remains the same. 
Consider the example in Figure~\ref{fig:running2}, given the branch 
if($x^9$$>$11) at $e_{10}$, the original mapping is $\theta(x^9)=12$. We allow  
a different value mapping, e.g., $\theta'(x^9)=13$, which retains the 
same truth value for the branch.
\item {\bf Intra-thread Order Constraints\ } The events in the sequence 
should follow the same order as in the original trace. More formally, 
we require that
$$
\begin{array}{rl}
\forall e_m, e_n \in prefix(\tau,e_i), s.t., & t^{e_m} = t^{e_n}. \\
 O(e_m) < O(e_n)\  & \Rightarrow O'(e_{m}) < O'(e_{n}) 
\end{array}
$$ 
This constraint specifies that two events from the same thread should 
follow the same order as reflected by the ids of the events.
\item {\bf Intra-thread Value Constraints\ } The value mapping of the 
variables should not contradict the constraints imposed by individual 
instruction.   Without loss of generality, suppose the instruction is 
a local assignment in the form of $x=y+z$, then we require that
$$
\bigwedge_{e_k \in prefix(\tau,e_i) \wedge inst^{e_k}=x=y+z}\
	\theta'(x)=\theta'(y)+\theta'(z)
$$
Recall that every variable may obtain a different value in the new 
trace $\tau'$ because it may read from some shared variable affected by 
the rescheduling. The constraint specifies that, the variables should 
be re-assigned to some values that are consistent with the instruction. 
Consider the example in Figure~\ref{fig:running2}, given the 
event $x^9=y^8+z^7$ at line 9, if the mappings $\theta'(y^8)$ and 
$\theta'(z^7)$ change to 3 and 10 respectively, the 
mapping $\theta'(x^9)$ should change to 13 to be consistent. 
%In the above, we use the local assignment event for demonstration, in general, the intra-thread value constraints should also be specified for the {\sf  heapr/heapw} events.
\end{itemize}

@@

To ensure feasibility under value relaxation
% of flow dependencies, 
we need to ensure the flow between read and write with some execution schedule. 
For example, in Figure \ref{fig:running2}, for the read at $e_8$ to read 
from the write t $e_1$, we need a schedule where 
$e_8$ happens after $e_1$ and other writes such as $e_3$ do not interleave them.

@@

An important concern that arises due to value relaxation 
%of flow dependencies 
is that heap accesses may change their meaning, i.e., they involve 
different base objects and no longer match with each other. As an 
illustration, we refer to Figure \ref{fig:heapAccess}. While the read at 
event $e_7$ appears to match the write at $e_5$, this is conditioned 
on the read at $e_6$ being linked to the assignment at $e_4$ (
object $o_2$ is the object denoted by $z$ according to the source code). 
Suppose event $e_6$ reads from $e_3$ due to the relaxation. Then $e_7$ and 
$e_5$ should not share the same base object. 
%Even worse, we do not know which events $e_7$ matches, because the base object is no longer known.

>>>>

\subsection{SSA Form of the Trace}~\label{sec:ssa}
First of all, we rewrite the trace into SSA form, such that every variable is
defined exactly once in the trace. This requirement is a prerequisite for 
the computation of the mapping, in which each variable is mapped to exactly 
one value.

\tool\ handles local assignments and heap accesses differently.
\begin{itemize}
\item {\bf Local Assignment\ } The SSA form for local assignments 
resembles the static SSA form in compiler optimizations, 
except that loops and recursions are fully resolved in the trace, 
obviating the need for {\sf Phi} nodes\footnote{While \tool\  in fact uses
Phi nodes in the second relaxation, for simplicity during discussion, we 
do not assume Phi nodes in this stage.} . More concretely, we replace the 
variable $v$ defined in an event, as well as the following uses of the 
definition, to a new variable $v^{id}$, where $id$ is the unique id 
of the event. The uniqueness of the id guarantees that no two events 
define the same variable, i.e., each variable is defined exactly once.  
In Figure~\ref{fig:running2}, for example,
the local variable $x$ defined at line 4 is renamed as $x^4$.
% where the line number is also used as event id,  the local variable $x$ defined at line 4 is renamed as $x^4$ in the trace, while $x$ defined at line 9 is renamed as $x^9$, which is used in the following two events.  At line 5, where the object $o3$ is created,  we define a reference variable $w^5$.  Internally,  the object $o3$ is encoded as a unique integer that indicates its heap address.
\item {\bf Heap accesses\ } The def/use relation between local heap accesses 
can be determined locally. Therefore, we treat them in the same way as 
local assignments. The def/use relations between shared heap accesses 
are more complex. We may link a read with different writes under different 
schedules, e.g., the read of {\tt y} at line 4 in Figure \ref{fig:running} can 
be from before line 1, line 1, or line 3. Therefore, we introduce two symbols to represent the shared read and shared write respectively and use them 
to reason about the possible options.
%establish different links flexibly. 
The details are as follows. 
\begin{itemize}
\item {\bf Local Heap Accesses\ } Given the write $x.f=y$ to a local heap 
location at event $e_{id}$,  we introduce a fresh local variable 
{\tt $l^{id}_{o.f}$} ($o$ is the runtime object referenced by $x$) to 
replace $x.f$. We also replace the uses of the location $o.f$ accordingly.
Here  the subscript $o.f$ is interpreted as a heap location, i.e., the 
field $f$ of the object $o$. Consider the example in 
Figure~\ref{fig:running2}. At lines 6-7, we replace the local heap 
accesses as the local variable $l^6_{o3.f}$.
%An assumption here is the uses and the definition still refer to the same base object $o$ in the predicted run, which is enforced through the constraint specification %(Section~\ref{sec:relax1}).
\item  {\bf Shared Heap Accesses\ }  Given the write $x.f=y$ or the read $y=x.f$ of a shared location at event $e_{id}$, we introduce two symbols, $W^{id}_{\tt o.f }$ and $R^{id}_{\tt  o.f }$ respectively, where $o$ is the runtime object referenced by $x$. Similarly, the subscript $o.f$ is interpreted as a heap location, i.e., the field $f$ of the object $o$. Note that in local heap encoding, 
both read and write are bounded to the same variable, whereas in shared
heap encoding, we use different variables to denote read and write, allowing
the solver to reason about the different dataflow.
\end{itemize}
We use the heap location such as $o.f$ in the symbolic form of heap accesses.
 An underlying assumption is that the heap location remains unchanged, 
which we ensure through additional constraints (Section~\ref{sec:relax1}).
\end{itemize}

@@

\begin{itemize}
\item {\bf Control Flow Constraints\ } The branches in the predicted run 
should take the same decisions as in the original trace so that each thread 
reproduces the same sequence of events\footnote{We allow path differences in
the second relaxation. This is to avoid presenting the encoding
of both relaxations together, which is overly complex.}.
%the encodings of both relaxation together, .  
%Specifically, we only need to reason about the branches prior to $e_i$.

Without loss of generality, we assume the branch event $e_k$ is in the form of $if(x<y)$. Then  we require that 
$$
\bigwedge_{e_k \in prefix(\tau,e_i) \wedge  inst^{e_k}=if(x<y).\ }  \theta'(x)<\theta'(y) 	\equiv \theta(x)<\theta(y)
$$ 
The constraint specifies that the branch condition should be evaluated as 
the same boolean value in the predicted run $\tau'$ as in the original 
trace $\tau$.
 Importantly, Unlike existing analyses \cite{yannis,pldi14}, we do not 
pose the requirement that the values flowing into branching statements 
remain the same, but adopt the relaxed requirement that 
the branch outcome remains the same.
%the evaluation 
%of branching expressions remains the same. 
Consider the example in Figure~\ref{fig:running2}, given the branch 
if($x^9$$>$11$) at $e_{10}$, the original mapping is $\theta(x^9)=12$. We allow  
a different value mapping, e.g., $\theta'(x^9)=13$, which retains the 
same truth value for the branch.
\item {\bf Intra-thread Order Constraints\ } The events in the sequence 
should follow the same order as in the original trace. More formally, 
we require that
$$
\begin{array}{rl}
\forall e_m, e_n \in prefix(\tau,e_i), s.t., & t^{e_m} = t^{e_n}. \\
 O(e_m) < O(e_n)\  & \Rightarrow O'(e_{m}) < O'(e_{n}) 
\end{array}
$$ 
This constraint specifies that two events from the same thread ($t^{e_m}$ denotes the thread of $e_m$) should 
follow the same order as in the original trace.
\item {\bf Intra-thread Value Constraints\ } The value mapping of the 
variables should not contradict the constraints imposed by individual 
instruction.   Without loss of generality, suppose the instruction is 
a local assignment in the form of $x=y+z$, then we require that
$$
\bigwedge_{e_k \in prefix(\tau,e_i) \wedge inst^{e_k}=x=y+z}\
	\theta'(x)=\theta'(y)+\theta'(z)
$$
Recall that every variable may obtain a different value in the new 
trace $\tau'$ because it may read from some shared variable affected by 
the rescheduling. The constraint specifies that, the variables should 
be re-assigned to some values that are consistent with the instruction. 
Consider the example in Figure~\ref{fig:running2}, given the 
event $x^9=y^8+z^7$ at line 9, if the mappings $\theta'(y^8)$ and 
$\theta'(z^7)$ change to 3 and 10 respectively, the 
mapping $\theta'(x^9)$ should change to 13 to be consistent. 
%In the above, we use the local assignment event for demonstration, in general, the intra-thread value constraints should also be specified for the {\sf  heapr/heapw} events.
\end{itemize}

@@

To ensure feasibility under value relaxation
% of flow dependencies, 
we need to ensure the flow between read and write with some execution schedule. 
For example, in Figure \ref{fig:running2}, for the read at $e_8$ to read 
from the write  $e_1$, we need a schedule where 
$e_8$ happens after $e_1$ and other writes such as $e_3$ do not interleave them.

@@

An important concern that arises due to value relaxation 
%of flow dependencies 
is that heap accesses may change their meaning, i.e., they involve 
different base objects and no longer match with each other. As an 
illustration, we refer to Figure \ref{fig:heapAccess}. While the read at 
event $e_7$ appears to match the write at $e_5$, this is conditioned 
on the read at $e_6$ being linked to the assignment at $e_4$ (
object $o_2$ is the object denoted by $z$ according to the source code). 
Suppose event $e_6$ reads from $e_3$ due to the relaxation. Then $e_7$ and 
$e_5$ should not match with each other as they do not share the same base object. 
%Even worse, we do not know which events $e_7$ matches, because the base object is no longer known.
